---
title: "Assignment 4, Task 1"
author: "Taylor Lockmann"
date: "3/6/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(naniar)

# For cluster analysis:
library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)
```
### Task 1: Agglomerative hierarchical clustering 

The following diagram explores similarities in stream chemistry in the Santa Barbara Coastal drainage area. 
```{r}
# Read in the data, clean up names, drop the timestamp column, and replace "-999" values with NA
stream_chemistry <- read_csv(here("data", "sbc_lter_registered_stream_chemistry.csv")) %>% 
  clean_names() %>% 
  select(-timestamp_local) %>% 
  replace_with_na_all(condition = ~.x == -999)

# Create a data frame with the mean of each site's metrics, keeping in instances of NAs
stream_means <- aggregate(stream_chemistry[, 2:11], list(stream_chemistry$site_code), mean, na.rm = TRUE)

# Drop rows where any of our measurement variables are missing
streams_complete <- stream_means %>% 
  drop_na()

# Scale the measurements
streams_scale <- streams_complete %>% 
  select(2:11) %>% 
  scale()

# Add site number as the row names
rownames(streams_scale) <- streams_complete$Group.1

# Compute Euclidian distances 
stream_euc_distance <- dist(streams_scale, method = "euclidian")

# View the output of euclidian distances:
#View(stream_euc_distance)

# Hierarchical clustering with complete linkage
streams_hc_complete <- hclust(stream_euc_distance, method = "complete")

# Make the plot:
ggdendrogram(streams_hc_complete,
             rotate = TRUE)+
  theme_minimal()+
  labs(x = "Site Code")
```

**Figure 1:** Dendrogram showing multivariate clustering for water chemistry by site. Analysis was performed using hierarchical clustering by complete linkage. From this diagram we can make the broad assumptions that sites MC00 and AB00 are the most similar sites, while site ON02 is the most dissimilar from the other sites.

**Source:** Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. Environmental Data Initiative. https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174. 

##### End Assignment 4, Task 1