---
title: "Assignment 4, Task 2"
author: "Taylor Lockmann"
date: "3/8/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(ggplot2)
library(nlstools)
library(here)
library(gt)
```

```{r}
# Data wrangling!


# Read in the data

fish_catch <- read_csv(here("data", "fish_catch.csv"))

fish_catch_1 <- fish_catch[-(1:4),] # Remove first four rows
  
fish_data <- fish_catch_1[-(64:67),] # Remove last four rows

fish_data <- fish_data %>% 
  rename(
    "year" = "World Wild Fish Catch and Farmed Fish Production, 1950-2012",
    "wild_catch" = "X2",
    "farmed_fish" = "X3",
    "total" = "X4"
  ) 
# Save the columns with numbers in them as numeric:
fish_data$total <- as.numeric(fish_data$total)
fish_data$wild_catch <- as.numeric(fish_data$wild_catch)
fish_data$farmed_fish <- as.numeric(fish_data$farmed_fish)
fish_data$year <- as.numeric(fish_data$year)

fish_data <- fish_data %>% 
  mutate(year_new = (year - 1950))
```

## 1. Exploratory graph of wild catch over time

```{r}
ggplot(data = fish_data, aes(x = year_new, y = wild_catch))+
  geom_point()+
  theme_minimal()+
  labs(
    x = "Year",
    y = "Wild Catch (millions of tons)"
  )
```

#### What type of relationship describes this trend?

It appears as if a logarithmic relationship describes this trend, leveling off at about year 40. The model for this relationship would be:

$P(t)=\frac{K}{1+Ae^{-kt}}$, where

- $K$ is the carrying capacity
- $A$ is $\frac{K-P_0}{P_0}$
- $k$ is the growth rate constant

```{r}
# Log transform the data we are evaluating, up to year 40, to estimate k
fish_data_exp <- fish_data %>% 
  filter(year_new < 40) %>% 
  mutate(ln_wild_catch = log(wild_catch))

fish_lm_k <- lm(ln_wild_catch ~ year_new, data = fish_data_exp)
fish_lm_k

# So it appears our coefficient (k) is approximately 0.04.
```

My initial estimate for *k* is 0.04. We can also estimate that *K* is close to 90 and *A* is 4.2.

## 2. Use Nonlinear Least Squares to find parameters.

Now that we have estimates for *k*, *K* and *A*, we'll use NLS to find the actual values.

```{r}
# Make our NLS model and run it
fish_nls <- nls(wild_catch ~ K/(1+A*exp(-r*year_new)),
                data = fish_data,
                start = list(K = 90, A = 4.2, r = 0.04),
                trace = TRUE)

# Format our model's outputs in tidy format
fish_model <- broom::tidy(fish_nls)
```

Our model with estimated parameters is:
$$P(t) = \frac{100.27}{1+4.316e^{-0.070t}}$$

```{r}
# Make a pretty table with our model outputs
fish_model %>% 
  gt() %>% 
  tab_header(
    title = "Outcome of Nonlinear Least Squares parameters analysis",
    subtitle = "Wild fish catch over time (all units in millions of tons)"
  ) %>% 
  fmt_number(
    columns = vars(estimate, std.error, statistic),
    decimals = 3
  ) %>% 
  cols_label(
    term = "Term",
    estimate = "Estimate",
    std.error = "Standard Error",
    statistic = "Statistic",
    p.value = "p-value"
  ) %>% 
  tab_options(
    table.width = pct(80)
  )
```

### 3. Combine original data and model output

```{r}
# Have our model make predictions at all the times in the original fish data
fish_predict <- predict(fish_nls)

# Bind predictions to original data frame:
fish_complete <- data.frame(fish_data, fish_predict)

# Plot together
ggplot(data = fish_complete, aes(x = year_new, y = wild_catch))+
  geom_point()+
  geom_line(aes(x = year_new, y = fish_predict))+
  theme_minimal()+
  labs(
    x = "Year",
    y = "Wild Catch (millions of tons)"
  )
```


**Source:** Global wild fish catch and aquaculture production, compiled by Earth Policy Institute with 1950-2010 from U.N. Food and Agriculture Organization (FAO), Global Capture Production and Global Aquaculture Production, electronic databases, at www.fao.org/fishery/topic/16140/en.



##### End Assignment 4, Task 2